Towards a classification framework for social machines

### 2013-02-26 latex/paper_new.tex
Abstract from dan

social machines, overview of what they are, their history
    people interacting with content and with each other

>>>>

The state of the art in human interaction with computation systems blurs the line between
computation performed by machine logic and algorithms, and those that result from input by
humans, arising from their own neurological processes and life experience. Current
socio-technical systems, known as `social machines' exploit the large-scale interaction of
humans with machines, motivated for numerous aims and purposes across the spectrums of
financial gain, charitable aid and simply for fun. In this paper we explore
the landscape of social machines, both past and present, towards the aim of defining an
initial classification framework. Through a number of knowledge elicitation
and refinement exercises we have identified the polyarchical relationship between
infrastructure, social machines and large-scale social initiatives. Our initial framework
describes classification constructs in the areas of {\it participation}, {\it participants} and
{\it motivation}. We present an initial classification of the largest social machines, as
an initial validation and demonstration of the use of our identified constructs.
There is rich potential for performing
analysis on the behaviour and phenomenology of social machines, and analysis of their
growth and outputs over time. Our future work is to elicit additional opinion,
classifications and validation from a wider audience, to produce a formal framework to
classify and aid in analysis of social machines.
